### 2023-10-06 main.tex
fix
c
Update main.tex
e

\begingroup
% \centering
\begin{center}
 \begingroup\LARGE
\bf Multispecies stirring process: duality and exact solution
\par\endgroup
 \vspace{3.5em}
 \begingroup\large \bf
% {Rouven Frassek}
 \par\endgroup
\vspace{2em}

@@

\section{The stirring process}

@@

The process studied in this paper is the {\em multi-species stirring process}. 

@@

particle of type $k$ at rate $\alpha_k^x >0$.

@@

The site-dynamics is instead of Glauber-type 
and, in the long-time limit, so-called non-equilibrium

@@

\vspace{.5cm}
\noindent
{\bf Remark:} in the following we will often consider type $N$ as a distinguished type
(called `holes'), whose number at each site is fixed by the numbers of particles 
of the remaining $N-1$ types (at the same site). It is interesting to observe that, in case we stop distinguishing the colors of the ``true'' particles of types $\{1,\ldots,N-1\}$, we retrieve the 
boundary-driven version of the partial symmetric exclusion process \cite{schutzSandow,carinci2013duality}.
The choice of type $N$ for the distinguished type is arbitrary and does not affect
the results. In particular, we shall see in Section \ref{sectionDuality} about duality, that
the dual process eventually has only particles of type $N$ on the graph.
Interpreting type $N$ as an hole, one can then say that the dual process voids the graph.

@@

    \Omega:=\bigtimes_{x\in V} \Omega_{x}

@@

We denote a particle configuration as $\bm{n}\in \Omega$, where $\bm{n}=(n_{a}^{x})_{x\in V,\,a\in\{1,\ldots,N\}}$ .\\
The dynamics consists in an exchange of particles (at Poissonian times) between connected vertices. Moreover, each vertex  exchanges particles with the external environement. The infinitesimal generator of the process reads

@@

where  $ \omega_{x,y}\geq 0$ are so-called conductances and $\Gamma_{x}\geq 0$ local inhomogeneities. The generator $\mathcal{L}_{x,y}$ is called the \textit{edge generator}, while $\mathcal{L}_{x}$ is called the \textit{site generator}. These linear operators act on functions $f:\Omega\to \mathbb{R}$ as follows

@@

1\qquad &\text{if}\quad y=x,\;b=a\\

@@

\end{cases}\,.
{\color{red} notation}
This means that on the edge $(x,y)\in \mathcal{E}$  a particle of type $a$ at site $x$ is exchanged with a particle of type $b$ at site $y$ at rate $\omega_{x,y}n_{a}^{x}n_{b}^{y}$. Furthermore, on each site $x\in V$  a particle of type $b$ is replaced with a particle of type $a$ at rate $\Gamma_{x}\alpha_{a}^{x}n_{b}^{x}$. For the sake of notation we denote $|\alpha^{x}|=\sum_{a=1}^{N}\alpha_{a}^{x}$. 
See Figure \ref{fig:1} and  Figure \ref{fig:2} for a visualisation of the dynamics. 
\begin{figure}[h]
    \centering
    \includegraphics[scale=0.45]{BulkStirring.eps}
    \caption{The edge dynamics. Maximal occupation is $\nu=6$. A particle of type $2$ is removed in $x$ and placed in $y$. A particle of type $3$ is removed in $y$ and placed in $x$. This transition occurs at rate $n_{2}^{x}n_{3}^{y}=2\cdot 3$.}
    \label{fig:1}
\end{figure}
\begin{figure}[h]
    \centering
    \includegraphics[scale=0.4]{BoundaryStirring.eps}
    \caption{The site dynamics. Here the maxiaml occupation is $\nu=6$. The reservoir, represented by a square "box" containing all types of particles, remove a type $2$ from site $x$ and replace it with a type $3$. This transition occurs at rate $\alpha_{3}^{x}n_{2}^{x}$.}
    \label{fig:2}
\end{figure}
\subsection{Reversible measures}
The process described by the generator \eqref{Generator} is reversible with respect to the homogeneus product measure \begin{equation}\label{reversibleMeasure}
\mu_{rev}=\bigtimes_{x\in V}\mu_{rev}^{x}

@@

\alpha_{a}^{x}=\alpha_{a}\qquad \forall x\in V\,.
This measure has marginals $\mu_{rev}^{x}$ given by 
 \mu^{x}_{rev}\sim \text{Multinomial}\left(\twoj,\rho_{1},\ldots,\rho_{N}\right)\quad \text{where}\quad \rho_{a}=\frac{\alpha_{a}}{|\alpha|}\co
namely
\mu_{rev}^{x}(n^{x})=\frac{\nu!}{\prod_{a=1}^{N}n_{a}^{x}!}\prod_{a=1}^{N}\rho_{a}^{n_{a}^{x}}
where we denote $|\alpha|=\sum_{a=1}^{N}\alpha_{a}$. 
This can be proved just by imposing the detailed balance conditions for the edge and for the site generators. If condition \eqref{reversibilityCondition} is not met, then reversibility is lost: each reservoir at site $x$ wants to fix its own density $\alpha^{x}=(\alpha_{1}^{x},\ldots,\alpha_{N}^{x})$ and currents will arise when at least two of them are different.

{\color{red}add/improve explanaition of entire subsection} {\color{blue}Is it not standard?}
\subsection{The Lie algebraic description}

@@

\\ \\
\textbf{Remark}: If we stop distinguishing the types of particles we retrieve the closed formula for correlations in the non-equilibrium steady state found in \cite{frassek2020eigenstates}. Indeed, introducing $q=\sum_{j=1}^{m}c_{i}$ in the colour-blind situation \eqref{exactCorrelations} becomes
	\langle n^{y_{1}}_{s_{1}}\cdots n^{y_{m}}_{s_{m}}\rangle_{\text{stat}}=&\sum_{q=0}^{m}(\alpha-\beta)^{q}\beta^{m-q}\frac{\Gamma(L+2-q)}{\Gamma(L+2)}\sum_{c_{1}=0}^{1}\ldots\sum_{c_{m}=0}^{1}\mathbbm{1}_{\{c_{1}+\cdots+c_{m}=q\}}
	\\&\prod_{k=1}^{m}\left(L+2-y_{k}-\sum_{i=k}^{m}c_{i}\right)^{c_{k}}\dt
We observe that, at most $q$ factors do survive in the product over $k$. For any fixed $q$ we must consider all possible values of $(c_{i})_{i\in\{1,\ldots,m\}}$ such that its sum is equal to $q$. For each of these $(c_{i})_{i\in\{1,\ldots,m\}}$ we add a term given by  
	\prod_{k=1}^{m}\left(L+2-y_{k}-\sum_{i=k}^{m}c_{i}\right)^{c_{k}}\dt
We can introduce $\ell_{1},\ldots,\ell_{q}\in\{1,\ldots,m\}$ such that $\ell_{1}<\ell_{2}<\cdots<\ell_{q}$ and write 
	\sum_{c_{1}=0}^{1}\ldots\sum_{c_{m}=0}^{1}\mathbbm{1}_{\{c_{1}+\cdots+c_{m}=q\}}\prod_{k=1}^{m}\left(L+2-y_{k}-\sum_{i=k}^{m}c_{i}\right)^{c_{k}}=\sum_{1\leq \ell_{1}<\ldots<\ell_{q}\leq m}\prod_{k=1}^{q}\left(L+2-y_{\ell_{k}}-(q-k+1)\right)\dt
Moreover, we observe that 
	\frac{\Gamma(L+2-q)}{\Gamma(2+L)}=\prod_{k=1}^{q}\frac{1}{(2+L-k)}\dt
All in all we obtain 
\langle n^{y_{1}}_{s_{1}}\cdots n^{y_{m}}_{s_{m}}\rangle_{\text{stat}}=\sum_{q=0}^{m}(\alpha-\beta)^{q}\beta^{m-q}\sum_{1\leq \ell_{1}<\ldots<\ell_{q}\leq m}\prod_{k=1}^{q}\frac{\left(L+1-y_{\ell_{k}}-q+k\right)}{(2+L-k)}\dt
that is exactly equation (4.26) of \cite{frassek2020eigenstates}.

>>>>

\begingroup
% \centering
\begin{center}
 \begingroup\LARGE
%\bf Multispecies stirring process: duality and exact solution
\bf Multispecies stirring process \\
 with open boundaries
\par\endgroup
 \vspace{3.5em}
 \begingroup\large \bf
% {Rouven Frassek}
 \par\endgroup
\vspace{2em}

@@

%
%
%
%
\section{The stirring process with open boundaries}

@@

The process studied in this paper is the {\em multi-species stirring process}
with open boundaries. 

@@

particle of type $a$ at rate $\alpha_a^x >0$.

@@

The site-dynamics is instead of Glauber-type. 
In the long-time limit, a so-called non-equilibrium

@@

\vspace{.5cm}
\noindent
{\bf Remark.} In the following we will often consider type $N$ as a distinguished type
(called `holes'), whose number at each site is fixed by the numbers of particles 
of the remaining $N-1$ types. It is interesting to observe that, in case we stop distinguishing the types of the ``true'' particles of types 
(those of types $\{1,\ldots,N-1\}$), we retrieve the 
boundary-driven version of the partial symmetric exclusion process \cite{schutzSandow,carinci2013duality}.
The choice of type $N$ for the distinguished type is arbitrary and does not affect
the results. In particular, we shall see in Section \ref{sectionDuality} about duality, that
the dual process eventually has only particles of type $N$ on the graph.
Interpreting type $N$ as an hole, one can then say that the dual process voids the graph.

@@

    \Omega:=\bigotimes_{x\in V} \Omega_{x}

@@

We denote a particle configuration on the graph as $\bm{n}\in \Omega$, where $\bm{n}=(n_{a}^{x})_{x\in V,\,a\in\{1,\ldots,N\}}$.
The infinitesimal generator of the process reads

@@

where  $ \omega_{x,y}\geq 0$ are so-called conductances and $\Gamma_{x}\geq 0$ are the couplings to reservoirs. 
The generator $\mathcal{L}_{x,y}$ is called the \textit{edge generator}, while $\mathcal{L}_{x}$ is called the \textit{site generator}. These linear operators act on functions $f:\Omega\to \mathbb{R}$ as follows

@@

1\qquad &\text{if}\quad y=x,\;b=a\;,\\

@@

Thus the dynamics consists in an exchange of particles between connected vertices at Poissonian times. 
This means that on the edge $(x,y)\in \mathcal{E}$  a particle of type $a$ at site $x$ is exchanged with a particle of type $b$ at site $y$ at rate $\omega_{x,y}n_{a}^{x}n_{b}^{y}$. 
Moreover, each vertex  exchanges particles with the external environment (reservoirs). 
Namely, on each site $x\in V$  a particle of type $b$ is replaced with a particle of type $a$ at rate $\Gamma_{x}\alpha_{a}^{x}n_{b}^{x}$. 
%
%For the sake of notation we denote $|\alpha^{x}|=\sum_{a=1}^{N}\alpha_{a}^{x}$. 
%See Figure \ref{fig:1} and  Figure \ref{fig:2} for a visualisation of the dynamics. 
%\begin{figure}[h]
%    \centering
%    \includegraphics[scale=0.45]{BulkStirring.eps}
%    \caption{The edge dynamics. Maximal occupation is $\nu=6$. A particle of type $2$ is removed in $x$ and placed in $y$. A particle of type $3$ is removed in $y$ and placed in $x$. This transition occurs at rate $n_{2}^{x}n_{3}^{y}=2\cdot 3$.}
%    \label{fig:1}
%\end{figure}
%\begin{figure}[h]
%    \centering
%    \includegraphics[scale=0.4]{BoundaryStirring.eps}
%    \caption{The site dynamics. Here the maxiaml occupation is $\nu=6$. The reservoir, represented by a square "box" containing all types of particles, remove a type $2$ from site $x$ and replace it with a type $3$. This transition occurs at rate $\alpha_{3}^{x}n_{2}^{x}$.}
%    \label{fig:2}
%\end{figure}
\subsection{Reversible measures (equilibrium)}
For a particular choice of the reservoir parameters one has an $N$-parameter family of reversible measures. More precisely
when the parameters are the same on each site, i.e.

@@

\alpha_{a}^{x}=\alpha_{a}\qquad \forall x\in V\,,
then the process described by the generator \eqref{Generator} is reversible with respect to the 
homogeneous product measure 
\label{reversibleMeasure}
\mu_{rev}=\bigotimes_{x\in V}\mu_{rev}^{x}
with marginals $\mu_{rev}^{x}$ given by 
 \mu^{x}_{rev}\sim \text{Multinomial}\left(\twoj,\rho_{1},\ldots,\rho_{N}\right)\,.
Here
\rho_{a}=\frac{\alpha_{a}}{|\alpha|}\co
is the density of specie $a$ and we used the notation $|\alpha|=\sum_{a=1}^{N}\alpha_{a}$. Explicitly, 
\mu_{rev}^{x}(n^{x})=\frac{\nu!}{\prod_{a=1}^{N}n_{a}^{x}!}\prod_{a=1}^{N}\rho_{a}^{n_{a}^{x}}\,.
This can be proved  by checking that detailed balance is satisfied. 
If condition \eqref{reversibilityCondition} is not met, then reversibility is lost: each reservoir at site $x$ wants to fix its own set of densities $\rho^{x}=(\rho_{1}^{x},\ldots,\rho_{N}^{x})$ with
\rho_{a}^x=\frac{\alpha_{a}}{|\alpha^x|}
and  $|\alpha^x|=\sum_{a=1}^{N}\alpha_{a}$. 
Currents will arise when at least two reservoirs have different densities.

\subsection{The Lie algebraic description}

@@
